Article Title

### 2014-01-06 plos.tex
complete Discussion

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neural populations have proven useful as descriptions of the functional connectivity in the embedding circuit.  Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.  What kinds of correlation effects dominate the population activity in cortical microcircuits? To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.  In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.  Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from sample correlations. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.

@@

\paragraph{Improved estimation performance by regularization}
Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.

Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimation of the partial correlations \cite{Schafer:2005}.  In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix} and these two concepts are rarely distinguished.  The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise. As a result, on average, the sample correlation matrix is far from the true correlation matrix.  In contrast, reduced estimates are typically less susceptible to sampling noise although can introduce their respective biases.

@@

First, we demonstrated that, in simulations with known true correlation structures, estimators that imposed a matching structure were generally more efficient than the other estimators.  The estimators’ relative performance was evaluated by cross-validation.   The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.  Caveats for this kind of inference are reviewed in the Discussion.

@@

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The icovariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$,  using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.

@@

Various other regularization schemes also produce improved estimates

We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic systems with linear dependencies, these correspond to graphical modes with distinct types of structure  (\figref{02}{row 1}).

@@

The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelih\emph{normal log-likelihood loss} defined as ood function $L(\Sigma|C_{\sf 0})$ as
\begin{equation}
    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
\end{equation}
Although the normal log-likelihood loss and the normal log-likelihood appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from the estimate $C$ to the truth $\Sigma$.

We normalized the loss function by $\frac 1 p$ to make its value comparable across different population sizes.

@@

\paragraph{Covariance estimation in neural data}
We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus. Using acousto-optic deflectors (AOD) to steer the laser in 3D, we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.  This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (See Methods). The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{03}{F and G}). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}). Overall, 31 imaged sites from 24 animals were included in the study.

@@

\section*{Discussion}
\paragraph{Functional connectivity}
How the anatomical organization of neural circuits gives rise to their function is a foremost question in neuroscience.  Effective phenomenological descriptions of neural activity, once discovered, enable investigation of underlying physiological mechanisms.  For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.

Similarly, the term \sq{functional connectivity} denotes a statistical summary of observed \emph{multineuronal} patterns of activity thought to reflect internal, recurrent connectivity in neural circuits.  In statistical terms, functional connectivity is an empirical estimate of a feature of the joint probability distribution of the neurons’ activity.  Although functional connectivity may not provide a detailed and unambiguous interpretation of underlying mechanisms, its utility in experimental neuroscience is that of a descriptive statistic to quantify the general organization of synchronous activations of neuronal populations.  Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons.  To remove the confounding effects of stimuli, the average stimulus responses are usually subtracted or otherwise accounted for.

Pairwise correlations (including cross-correlograms and joint peristimulus histograms) between the spiking activities of neurons boast a rich history as a measure of functional connectivity.  Interpreted as either the statistical signatures of synaptic connectivity \cite{Gerstein:1964, Perkel:1969, Denman:2013} or of Hebbian neuronal assemblies \cite{Gerstein:1989, Lopes:2013}, pairwise correlations have been instrumental in discovering various principles of the functional organization of neural circuits, as only partially enumerated in the Introduction.  Several other measures of functions connectivity have emerged including probabilistic models

Parsimony is a desired attribute of a measure of functional connectivity: insignificant or contingent detail.

\paragraph{Temporal scales of functional connectivity}
Analyses of functional connectivity depend profoundly on their temporal resolution.  At millisecond-scale resolution, cross-correlograms of the spiking activity of pairs of cells manifest statistical signatures of monosynaptic connections \cite{Gerstein:1964,  Perkel:1969, Moore:1970, Gerstein:1989, Alonso:1998, Denman:2013}.  Other methods require temporal resolution on the order of membrane time constants (milliseconds to tens of milliseconds): \emph{e.g.}~statistical models that mimic membrane integration (\cite{Pillow:2008}) and information-theoretic and statistical physics models that bin spike trains finely to ensure a binary representation \cite{Schneidman:2006, Tkacik:2006, Tkacik:2013}. At slower scales of tens to hundreds milliseconds, temporal sequencing of spikes can no longer resolve directions of monosynaptic interactions and any expression of functional connectivity summarizes chains of synaptic transmissions with more complex, emergent properties of network computation: the analysis shifts from to detecting co-active cell assemblies and the low-rank structure of correlations \cite{Ch:2010, Lopes:2011, Lopes:2013}.

The temporal resolution of calcium signals is limited by calcium dye kinetics. Fast imaging techniques combined with advanced spike inference algorithms have been demonstrated to provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010, Katona:2012, Cotton:2013}. However, high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.  Therefore, functional connectivity studies from calcium signals have focused on the slower temporal scales (tens or hundreds of milliseconds). In this study, too, we chose 20 Hz for the temporal precision of spike inference and bin widths of 150 ms for computing correlations.

\paragraph{Estimation of functional connectivity depends on recorded population size and density}
Analyses of functional connectivity depend profoundly on the sizes and densities the recorded population.  Pairwise recordings may contain statistical signatures of monosynaptic connections but conditioning on the activity of all the other cells may help disambiguate the network of interactions in the circuit.  This distinction has been described as \emph{effective connectivity}: the best guess at the most direct effects exerted by neurons on each other \cite{Gerstein:1989}.  For example, the interaction terms in spatiotemporal generalized linear models \cite{Pillow:2008} or Ising-like maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Ganmor:2011} express the conditional dependencies between neurons, which only have weak relationship with the pairwise statistics to which they are fitted.  If neurons can be assumed to exert nearly linear effects on each other, then conditional dependencies between neurons are expressed by the partial correlations.  For effectual computation of conditional dependencies, the recording must include a sufficient fraction of the neurons in the embedding circuit, or even the complete population. Otherwise, interactions with the unobserved portions of the circuit will introduce false conditional dependencies between the observed neurons. For this reason, the most successful applications of statistical models of population activity thus far have been in in vitro preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible.

When inserted in cortical tissue, electrode arrays record from a small fraction of cells in a given volume.  Perhaps due to such sparse sampling of neurons, the structure of partial correlations has not, to our knowledge, been used to describe the functional connectivity in multielectrode recordings.  Two-photon calcium imaging allows simultaneous measurements of the activity of nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  Therefore, two-photon imaging opens opportunities for computing the structure of conditional dependencies between neurons with arguably closer correspondence to the underlying physiological interactions.

\paragraph{Finding improved representation of the correlation structure}
We began this study from the somewhat simplistic hypothesis that the pairwise partial correlations in local cortical microcircuits should correspond more closely to the synaptic or effective connectivity than the sample correlations.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we required that the partial correlations be sparse. To account for interactions with the rest of the brain, we applied the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.  A priori, the validity of this assumption was debatable:  On the temporal scales of 150 ms, the pairwise effects between neurons could fade in prominence compared to the co-activation of larger cell assemblies and global circuit fluctuations, which would be poorly fitted by sparse partial correlations.  In addition, strong nonlinearities in interactions between neurons could disrupt the relationship between conditional independence and absence of partial correlation.

To verify our hypothesis, we compared the effect of imposing this structure on the estimation of the noise covariance matrices relative to estimations based on other assumptions.  The finding that the sparse+latent structure of correlations led to better estimation than the other estimators lends a degree of credence to its structural representation of the circuit’s functional connectivity.  Of course, this finding could be overturned or augmented by finding an even more efficient covariance estimator for these types of recordings and we welcome the challenge.

In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.  This does not limit the applicability of its conclusion to normal distributions. Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).  Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.  If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of functional connectivity.

>>>>

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neural populations have proven useful as descriptions of the functional connectivity in the embedding circuit.  Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.  What kinds of correlation effects dominate the population activity in cortical microcircuits? To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.  In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.  Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from sample correlations. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.

@@

\paragraph{Improving estimation through regularization}
Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.

Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.

In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix}, which is rarely distinguished from the true correlation matrix.  The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise. As a result, on average, the sample correlation matrix is far from the true correlation matrix.  In contrast, reduced estimates are typically less susceptible to sampling noise although can introduce their respective biases.

@@

First, we demonstrated that, in simulations with known true correlation structures, estimators that imposed a matching structure were generally more efficient than the other estimators.  The estimators’ relative performance was evaluated by cross-validation.   The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.  Caveats for this type of inference are reviewed in the Discussion.

@@

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The icovariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.  In the case of noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.

@@

We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{row 1}).

@@

This loss function relates to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ through the following expression:
\begin{equation}
    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
\end{equation}
Although the two functions appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from an arbitrary estimate $C$ to the given truth $\Sigma$.

The normalization by $\frac 1 p$ is not standard but serves to make the values of the loss functions comparable across different population sizes.

@@

\paragraph{Covariance estimation in neural data}
We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E}) in the visual cortex on the contralateral side to the stimulus. Using acousto-optic deflectors (AOD) to steer the laser in 3D, we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.  This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (See Methods). The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{03}{F and G}). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}). Overall, 31 imaged sites from 24 animals were included in the study.

@@

\section*{Discussion}
\paragraph{Functional connectivity}
Effective phenomenological or statistical descriptions of neural activity can serve as a starting point for investigation of underlying physiological mechanisms.  For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity and principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.  The search for similarly effective descriptions of observed \emph{multineuronal} activity patterns that are not reducible to the response properties of the individual cells has produced various statistical expressions collectively known as \sq{functional connectivity}.  Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons and is thought to reflect internal local recurrent connectivity, shared interactions with other regions, or endogenous network activity.  To remove the confounding effects of stimuli, the average stimulus responses are usually subtracted or otherwise accounted for. Although various measures of functional connectivity are only phenomenological descriptions of population activity, they provide experimental neuroscientists with tools to relate population activity to the anatomical and functional architecture of neural circuits and to generate hypotheses about underlying mechanisms.

\paragraph{Functional connectivity as a network of pairwise interactions}
The intended purpose of many studies of functional connectivity is to estimate the circuit’s anatomical connectivity solely from the observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are still ambiguous as they can arise from diverse network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from a larger subset of neurons in the circuit, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Since direct causal interactions between neurons produce statistical dependency even after conditioning on the state of the rest of the network, conditionally independent neurons can be inferred to lack a direct causal influence.  Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to include biophysically plausible synaptic integration and membrane kinetics, individual neurons’ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.  Another family of models are maximum entropy models that are constrained by the observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2009, Shlens:2009} .  Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with fewest other assumptions about the distribution: the investigator need only choose which statistics to match.  Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \emph{i.e.}\;conditional dependencies between pairs of neurons.

The inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the interaction between a pair of neurons on the activity of the other neurons under a given probabilistic model, the recording must include all the other neurons with which the pair can potentially interact. Otherwise, interactions with the unobserved portions of the circuit will introduce false conditional dependencies between the observed neurons. For this reason, the most successful applications of statistical models of population activity thus far have been in in vitro preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible \cite{Pillow:2008}. When inserted in cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the conditional dependencies between pairs of cells. Perhaps for this reason, until our study, partial correlations have not been used to describe the functional connectivity between cortical neurons.  Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.

Because the validity of the inference of interactions from probabilistic models depends on both the correctness of their mathematical form in the model and the completeness of the recorded population, it is far from clear which approach provides best correspondence to the anatomical connectivity in the circuit, with no experimental evidence yet available to address this question.  The result is often treated as a graphical representation of the functional connectivity without a statement about its anatomical implications.  Whatever these implications may be, investigators have applied graph-theoretical analyses to the inferred networks to make statements about the design principles and efficiency of circuit architectures \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity despite some of the disadvantages.  The temporal resolution of calcium signals is limited by calcium dye kinetics. Fast imaging techniques combined with advanced spike inference algorithms have been demonstrated to provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010, Katona:2012, Cotton:2013}. However, such high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.  For what calcium imaging lack in temporal resolution, it make up in spatial resolution and the ability to characterize the spatial arrangements of cells and identify their types.  Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  Such completeness may open new opportunities for inferring networks of conditional dependencies between neurons with greater validity than with electrophysiological recordings.

The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and maximum entropy models on the binary domain).  Most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .  Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without a statement about underlying mechanisms.

In our study, we sought to express pairwise connectivity networks among neurons in local cortical microcircuits. We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions in with sparse partial correlations.  The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.  The interactions (conditional dependencies) between pairs of neurons inferred by this distributed are expressed by the precision matrix (inverse of the covariance) or, equivalently, by the matrix of the pairwise partial correlations. Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than sample correlations.

\paragraph{Functional connectivity as coactivations}
Another approach to describe the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}. Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}. Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.  Coactivation patterns could also be explained as shared input from unobserved parts of the circuit.  Coactivations could also be explained by global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.

Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but analysis of coactivations method shifts the focus from detailed interactions to the collective behavior.

In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.

\paragraph{Combining pairwise interactions and coactivations}
In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.

In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.  The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.
\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.  The problem becomes increasingly worse as the size of the recorded population of neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As the recorded neuronal population sizes increased gradually, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption are observed\cite{Tkacik:2013}.  However, with ambitious projects, such as the BRAIN initiative currently underway \cite{Alivisatos:2013} aiming to record from significantly larger populations, the problem must be addressed by regularizing the solution. Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data. However, limiting the solution to an inappropriate subspace does not allow significant improvement.

Several strategies have been developed to limit the model space in order to improve the quality of the estimate. For example, \cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina. Generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.

In our study, regularization was also accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions.

Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in big neural data will entail the search for optimal regularizations that may involve combinations of heuristic rules and numerical techniques specifically designed for each type of neural circuit.

\paragraph{Model selection}
In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.  However, this does not limit the applicability of its conclusions to normal distributions. Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).  Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.  If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, a simple model can produce a better estimate even if it does not correctly represent the real dependencies in the data than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.  This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement.

\paragraph{Physiological interpretation and future directions}
Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, alludes synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred interaction network was substantially different from that inferred from by considering the network of most significant sample correlations (\figref{05}{F, H, and I}).  In particular, sample correlations do not draw attention to the absence of correlation between a pair of neurons that should be strongly correlated based on their common correlations with other neurons.  The $C_{\sf sparse+latent}$ estimator identified many such pairs and inferred negative interactions between them (\figref{05}{F}).  The resulting network had a larger fraction of negative interactions than the fraction of negative correlations in the sample correlation matrix.  The negative interactions had a distinct spatial organization from   Reproducing our study with labeled interneurons of various types could indicate whether the negative interactions
